---
title: "Data tidying code"
author: "Bernard, Estrada, Finkel, Marshall & Nelaballi"
date: "1 June, 2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

##### Install packages

```{r setup, include=FALSE}

library(tidyr)
library(dplyr)
library(plyr)
library(lubridate)
library(pander) # pretty table
library(gridExtra) # par(mfrow) equivalent for ggplot
library(lme4) # for glmer for neg binom
library(bbmle)
library(heplots)  # for the paired plots
library(sjPlot)
library(latticeExtra)

```

##### Setting working directory and reading in files

```{r, echo=TRUE}

# setting WD
setwd("D:/RESEARCH/leeches")

# read in the leech data
d <- read.csv("Complete Leech data entry sheet_2020-09-30.csv")

# read in the altitude data
altitude <- read.csv("GP_altitude.csv")


```

##### Tidying leech data - removing unwanted rows and columns

```{r, echo=TRUE}

head(d) 

# first row should be removed 
d <- read.csv("Complete Leech data entry sheet_2020-09-30.csv", skip = 1, header=TRUE, stringsAsFactors=T)
head(d)

# drop valueless columns
d <- select(d, -X, -X.1, -X.2, -X.3, -X.4, -X.5)

# drop valueless rows - where nothing was entered/or incorrectly entered for the response variable
length(d$Count.pacat.biasa)
length(d$Count.pacat.tupai)
levels(d$Count.pacat.biasa)
levels(d$Count.pacat.tupai)
d <- filter(d, d$Count.pacat.tupai != "")
d <- filter(d, d$Count.pacat.tupai != "?")
d <- filter(d, d$Count.pacat.biasa != "?")
d <- filter(d, d$Count.pacat.biasa != "")
d <- filter(d, d$Count.pacat.biasa != "1-3 jam")
levels(d$Count.pacat.biasa) <- gsub("tidak ada", "0", levels(d$Count.pacat.biasa)) #replacing one instance of "tidak ada" with "0"
levels(d$Count.pacat.tupai) <- gsub("tidak ada", "0", levels(d$Count.pacat.tupai)) #replacing one instance of "tidak ada" with "0"

# checking for unused levels & then dropping them
levels(d$Count.pacat.biasa)
levels(d$Count.pacat.tupai)
d$Count.pacat.biasa <- factor(d$Count.pacat.biasa) # dropping unused levels
d$Count.pacat.tupai <- factor(d$Count.pacat.tupai) # dropping unused levels
length(d$Count.pacat.biasa)
length(d$Count.pacat.tupai)
levels(d$Count.pacat.biasa)
levels(d$Count.pacat.tupai)

```

##### Tidying leech data - renaming variables

```{r, echo=TRUE}

colnames(d)
names(d)[names(d)=="Census.."] <- "Census"
names(d)[names(d)=="Segment.number"] <- "Segment"
names(d)[names(d)=="Hab"] <- "FT"
names(d)[names(d)=="Mulai"] <- "Start"
names(d)[names(d)=="Segment.length....tags."] <- "Segment_Tags"
names(d)[names(d)=="Distance..m."] <- "Distance"
names(d)[names(d)=="Tanggal"] <- "Date"
names(d)[names(d)=="Siapa.buat.sensus"] <- "Observer"
names(d)[names(d)=="Jam"] <- "Hour"
names(d)[names(d)=="Tingkat.lembab"] <- "Humidity_Census"
names(d)[names(d)=="Histori.hujan"] <- "Rain_History"
names(d)[names(d)=="Tipe.hujan"] <- "Rain_Type"
names(d)[names(d)=="Count.pacat.tupai"] <- "H.picta"
names(d)[names(d)=="Count.pacat.biasa"] <- "H.zeylanica"
names(d)[names(d)=="Tingkat.lembab.1"] <- "Humidity_Segment"
names(d)[names(d)=="Cuaca"] <- "Weather"
names(d)[names(d)=="Lama.hujan"] <- "Long_Rain"

```

##### Tidying leech data - translate levels

```{r, echo=TRUE}

# Humidity_Census

# original levels
levels(d$Humidity_Census)
summary(d$Humidity_Census)

# translate levels
d$Humidity_Census <- mapvalues(d$Humidity_Census, 
          from = c("", 
                   "basah",
                   "Basah",
                   "basah/lembab",
                   "kering",
                   "Kering",
                   "Kering ",
                   "kosong", #dry
                   "lambab",
                   "lemaba",
                   "lembab",
                   "Lembab",
                   "lembab/kering",
                   "lembabkosongtidak ada hujan"), #dry
          to = c("NA", 
                   "wet",
                   "wet",
                   "wet/moist",
                   "dry",
                   "dry",
                   "dry",
                   "dry",
                   "moist",
                   "moist",
                   "moist",
                   "moist",
                   "moist/dry",
                   "dry"))

# check new levels
levels(d$Humidity_Census)
summary(d$Humidity_Census)

# Rain_History

# original levels
levels(d$Rain_History)
summary(d$Rain_History)

# translate levels
d$Rain_History <- mapvalues(d$Rain_History, 
          from = c("",
                   ";agi hujan",
                   "kamarin",
                   "kamarin dulu",
                   "kamerin",
                   "kemarin",
                   "Kemarin",
                   "kemarin ada",
                   "kemarin ada hujan",
                   "Kemarin ada hujan",
                   "kemarin dulu",
                   "Kemarin dulu",
                   "kemarin duluh",
                   "kermarin",
                   "kermarin dulu",
                   "kosong",
                   "Kosong",
                   "lagi hujan",
                   "Lagi hujan",
                   "lagi hujan gerimis",
                   "lagi hujankuat",
                   "lemarin",
                   "tadi  malam",
                   "tadi malam",
                   "Tadi malam",
                   "tadi malam ",
                   "Tadi malam ada hujan",
                   "tadi malam kwat",
                   "tadi malam sedang" ), 
          to = c("NA",
                   "raining",
                   "yesterday",
                   "day before yesterday",
                   "yesterday",
                   "yesterday",
                   "yesterday",
                   "yesterday",
                   "yesterday",
                   "yesterday",
                   "day before yesterday",
                   "day before yesterday",
                   "day before yesterday",
                   "yesterday",
                   "day before yesterday",
                   "<48 hours ago",
                   "<48 hours ago",
                   "raining",
                   "raining",
                   "raining",
                   "raining",
                   "yesterday",
                   "last night",
                   "last night",
                   "last night",
                   "last night",
                   "last night",
                   "last night",
                   "last night" ))

# check new levels
levels(d$Rain_History)
summary(d$Rain_History)
d$Rain_History <- factor(d$Rain_History) # drop unused levels

# Rain_Type

# original levels
levels(d$Rain_Type)
summary(d$Rain_Type)

# translate levels
d$Rain_Type <- mapvalues(d$Rain_Type, 
          from = c("",
                   "gerimis",
                   "Gerimis",
                   "kuat",
                   "Kuat",
                   "kwat",
                   "Kwat",
                   "sedang",
                   "Sedang",
                   "tidak ad hujan",
                   "tidak ada",
                   "Tidak ada",
                   "tidak ada hujan",
                   "Tidak ada hujan"),
          to = c("NA",
                   "drizzle",
                   "drizzle",
                   "heavy",
                   "heavy",
                   "heavy",
                   "heavy",
                   "medium",
                   "medium",
                   "no rain",
                   "no rain",
                   "no rain",
                   "no rain",
                   "no rain"))
# check new levels
levels(d$Rain_Type)
summary(d$Rain_Type)

# Soil moisture

# original levels
levels(d$Humidity_Segment)
summary(d$Humidity_Segment)

# translate levels
d$Humidity_Segment <- mapvalues(d$Humidity_Segment, 
          from = c("",
                   "basag",
                   "basah",
                   "Basah",
                   "basah ",
                   "basash",
                   "bassah",
                   "biasah",
                   "cerah",  #translates to clear, coded as NA
                   "cerah ", #translates to clear, coded as NA
                   "gerimis", #coded as wet
                   "hujan",
                   "kemarin", #translates to yesterday, coded as NA
                   "kerbut", # perhaps kering, coded as dry
                   "kering",
                   "Kering",
                   "kering ",
                   "kernig",
                   "lagi hujan",
                   "lambab",
                   "Lambab",
                   "lembab",
                   "Lembab",
                   "lembab ",
                   "mendung", #translates to cloudy, coded as NA
                   "Mendung", #translates to cloudy, coded as NA
                   "mendung ", #translates to cloudy, coded as NA
                   "tidak ada" ), #coded as NA
          to = c("NA",
                   "wet",
                   "wet",
                   "wet",
                   "wet",
                   "wet",
                   "wet",
                   "wet",
                   "NA", 
                   "NA", 
                   "wet",
                   "wet",
                   "NA", 
                   "dry",
                   "dry",
                   "dry",
                   "dry",
                   "dry",
                   "wet",
                   "moist",
                   "moist",
                   "moist",
                   "moist",
                   "moist",
                   "NA",
                   "NA", 
                   "NA", 
                   "NA" ))
# check new levels
levels(d$Humidity_Segment)
summary(d$Humidity_Segment)

# Weather

# original levels
levels(d$Weather)
summary(d$Weather)

# translate levels
d$Weather <- mapvalues(d$Weather, 
          from = c("",
                   "basah", #coded as NA
                   "Basah", #coded as NA
                   "berkabut",
                   "cearh",
                   "cerah",
                   "Cerah",
                   "cerah ",
                   "cerah dan berangin",
                   "cerah/berangin",
                   "cerah/berkabut",
                   "cerah/berkabut asap",
                   "cerah/kabut asap tebal",
                   "cerah/mendung", #coded as NA
                   "cereah",
                   "clear",
                   "cloudy",
                   "cloudy drizzling little",
                   "cloudy/windy",
                   "cuaca", #coded as NA
                   "gelap", #coded as NA
                   "gerimis",
                   "Gerimis",
                   "gerimis kecil",
                   "gerimmis",
                   "hujan",
                   "Hujan",
                   "hujan gerimis kecil",
                   "hujan kwat",
                   "hujan sedang",
                   "kabut",
                   "Kabut",
                   "kabut asap",
                   "kebut",
                   "kering", #coded as clear
                   "Kering" , #coded as clear
                   "lembab", #coded as NA
                   "Lembab", #coded as NA
                   "mending",
                   "mendug",
                   "mendung",
                   "Mendung",
                   "mendung ",
                   "mendung berangin",
                   "mendung dan kabut",
                   "mendung gerimis",
                   "mendung gerimis kecil kecil",
                   "mendung, kabut dan gerimis", #coded as drizzle
                   "mendung/ gerimis",
                   "mendung/berkabut",
                   "mendung/cerah", #coded as NA
                   "mendung/gerimis",
                   "mendung/hujan",
                   "mendung/kabut",
                   "mengung",
                   "menung",
                   "tidak ada"), #coded as NA
          to = c("NA",
                   "NA",
                   "NA",
                   "fog",
                   "clear",
                   "clear",
                   "clear",
                   "clear",
                   "clear",
                   "clear",
                   "clear/fog",
                   "clear/fog",
                   "clear/fog",
                   "NA",
                   "clear",
                   "clear",
                   "cloudy",
                   "drizzle",
                   "cloudy",
                   "NA",
                   "NA",
                   "drizzle",
                   "drizzle",
                   "drizzle",
                   "drizzle",
                   "rain",
                   "rain",
                   "drizzle",
                   "rain",
                   "rain",
                   "fog",
                   "fog",
                   "fog",
                   "fog",
                   "clear",
                   "clear" ,
                   "NA",
                   "NA",
                   "cloudy",
                   "cloudy",
                   "cloudy",
                   "cloudy",
                   "cloudy",
                   "cloudy",
                   "cloudy/fog",
                   "drizzle",
                   "drizzle",
                   "drizzle",
                   "drizzle",
                   "cloudy/fog",
                   "NA",
                   "drizzle",
                   "rain",
                   "cloudy/fog",
                   "cloudy",
                   "cloudy",
                   "NA"))
# check new levels
levels(d$Weather)
summary(d$Weather)

```

##### Tidying leech data - order levels

```{r, echo=TRUE}

# order FT 

# original levels
levels(d$FT)
summary(d$FT)

# order levels
d$FT <- mapvalues(d$FT, 
          from = c("",
                   "AB",
                   "FS",
                   "LG",
                   "LS",
                   "MO",
                   "PS",
                   "UG"), 
          to = c("NA",
                 "3.AB",
                 "2.FS",
                 "5.LG",
                 "4.LS",
                 "7.MO",
                 "1.PS",
                 "6.UG"))

# check new levels
levels(d$FT)
summary(d$FT)
d$FT <- factor(d$FT)# drop unused levels
summary(d$FT)

# order Soil moisture

# original levels
levels(d$Humidity_Segment)
summary(d$Humidity_Segment)

# order levels
d$Humidity_Segment <- mapvalues(d$Humidity_Segment, 
          from = c("NA",
                   "wet",
                   "dry",
                   "moist"), 
          to = c("NA",
                 "3.wet",
                "1.dry",
                "2.moist"))

# check new levels
levels(d$Humidity_Segment)
summary(d$Humidity_Segment)

# order Weather

# original levels
levels(d$Weather)
summary(d$Weather)

# order levels
d$Weather <- mapvalues(d$Weather, 
          from = c("NA",
                   "fog",
                   "clear",
                   "clear/fog",
                   "cloudy",
                   "drizzle",
                   "rain",
                   "cloudy/fog"), 
          to = c("NA",
                   "3.fog",
                   "1.clear",
                   "2.clear/fog",
                   "5.cloudy",
                   "6.drizzle",
                   "7.rain",
                   "4.cloudy/fog"))

# check new levels
levels(d$Weather)
summary(d$Weather)

# order Rain_History

# original levels
levels(d$Rain_History)
summary(d$Rain_History)

# order levels
d$Rain_History <- mapvalues(d$Rain_History, 
          from = c("NA",
                   "yesterday",
                   "day before yesterday",
                   "<48 hours ago",
                   "raining",
                   "last night"), 
          to = c("NA",
                   "3.yesterday",
                   "2.day_before_yesterday",
                   "1.over_48_hours",
                   "5.raining",
                   "4.last_night"))

# check new levels
levels(d$Rain_History)
summary(d$Rain_History)

# order Rain_Type

# original levels
levels(d$Rain_Type)
summary(d$Rain_Type)

# order levels
d$Rain_Type <- mapvalues(d$Rain_Type, 
          from = c("NA",
                   "drizzle",
                   "heavy",
                   "medium",
                   "no rain"), 
          to = c("NA",
                   "2.drizzle",
                   "4.heavy",
                   "3.medium",
                   "1.no_rain"))

# check new levels
levels(d$Rain_Type)
summary(d$Rain_Type)

# drop level NA
d <- filter(d, d$Weather!= "NA")
d <- filter(d, d$Rain_Type!= "NA")
d <- filter(d, d$Humidity_Segment!= "NA")
d <- filter(d, d$Rain_History!= "NA")
d <- filter(d, d$Humidity_Census!= "NA")
d$Rain_Type <- factor(d$Rain_Type)
d$Humidity_Segment <- factor(d$Humidity_Segment)
d$Weather <- factor(d$Weather)
d$Rain_History <- factor(d$Rain_History)
d$Humidity_Census <- factor(d$Humidity_Census)

summary(d$Rain_Type)
summary(d$Weather)
summary(d$Humidity_Segment)
summary(d$Rain_History)
summary(d$Humidity_Census)

```

##### Tidying leech data - fix observer column

```{r, echo=TRUE}

# Observer

levels(d$Observer)
str(d$Observer)
summary(d$Observer)

# all instances with more than 1 obs assigned "> 1 obs", rest retain the value of d$Observer

d$Observer.S <- as.character(d$Observer)
d$Observer.S[nchar(d$Observer.S) > 2] <- "> 1 Obs"
d$Observer.S <- as.factor(d$Observer.S)

# check new levels
summary(d$Observer.S)

```

##### Tidying leech data - date & time index

```{r, echo=TRUE}

# Date

str(d$Date)

# extract year
d$year <- substring(d$Date,1,4)
d$year <- as.factor(d$year)
str(d$year)
levels(d$year)
levels(d$year) <- gsub("^$", "NA", levels(d$year)) # replacing level "" with "NA"
levels(d$year)

# time index
head(d$Date)

# convet to numeric dates
levels(d$Date) <- gsub("^$", "NA", levels(d$Date)) # replacing "" with "NA"
d$dates <- as.numeric(gsub("\\-", "", d$Date))
head(d$dates)

# to set 1st date as "1"
day.offset <- (range(d$dates)[1])-1
d$dates <- (d$dates-day.offset)
  
```

##### Tidying leech data - Misc. small fixes

```{r, echo=TRUE}

#14B-09 labeled UG, should be LG (site partition LG.II)
d$FT[d$Segment=='14B-09'] <- '5.LG'

#13A-01, 13B-08 labeled UG, should be LG
d$FT[d$Segment=='13A-01'] <- '5.LG'
d$FT[d$Segment=='13B-08'] <- '5.LG'

#7A-09, 7B-06, switches between MO.I MO.II within a segment, but more in MO.II so relabel?
d$FT[d$Segment=='7A-09'] <- '7.MO'
d$FT[d$Segment=='7B-06'] <- '7.MO'

```

##### adding altitude data to leech data

```{r, echo=TRUE}

d <- merge(d, altitude, by.x = "FT", by.y = "ï..FT", all.x = TRUE)

```

##### Assign site partitions to leech file based on segment / forest type

```{r, echo=TRUE}

colnames(d)

#PS
d$partition[d$FT == "1.PS"] <- as.character("PS.I")

#FS
d$partition[d$FT == "2.FS"] <- as.character("FS.I")

#AB
d$partition[d$Segment == "3A-02"] <- as.character("AB.I")
d$partition[d$Segment == "3A-03"] <- as.character("AB.I")
d$partition[d$Segment == "3A-06"] <- as.character("AB.I")
d$partition[d$Segment == "3A-07"] <- as.character("AB.I")
d$partition[d$Segment == "3A-10"] <- as.character("AB.I")
d$partition[d$Segment == "3B-01"] <- as.character("AB.I")
d$partition[d$Segment == "3B-04"] <- as.character("AB.I")
d$partition[d$Segment == "3B-05"] <- as.character("AB.I")
d$partition[d$Segment == "3B-08"] <- as.character("AB.I")
d$partition[d$Segment == "3B-09"] <- as.character("AB.I")

d$partition[d$Segment == "11A-08"] <- as.character("AB.I")
d$partition[d$Segment == "11B-01"] <- as.character("AB.I")

d$partition[d$Segment == "4A-01"] <- as.character("AB.II")
d$partition[d$Segment == "4A-02"] <- as.character("AB.II")
d$partition[d$Segment == "4A-03"] <- as.character("AB.II")
d$partition[d$Segment == "4B-06"] <- as.character("AB.II")
d$partition[d$Segment == "4B-07"] <- as.character("AB.II")
d$partition[d$Segment == "4B-08"] <- as.character("AB.II")

d$partition[d$Segment == "12A-01"] <- as.character("AB.II")
d$partition[d$Segment == "12A-02"] <- as.character("AB.II")
d$partition[d$Segment == "12A-03"] <- as.character("AB.II")
d$partition[d$Segment == "12A-07"] <- as.character("AB.II")
d$partition[d$Segment == "12A-08"] <- as.character("AB.II")
d$partition[d$Segment == "12B-01"] <- as.character("AB.II")
d$partition[d$Segment == "12B-02"] <- as.character("AB.II")
d$partition[d$Segment == "12B-06"] <- as.character("AB.II")
d$partition[d$Segment == "12B-07"] <- as.character("AB.II")
d$partition[d$Segment == "12B-08"] <- as.character("AB.II")

#LS
d$partition[d$Segment == "5A-01"] <- as.character("LS.I")
d$partition[d$Segment == "5A-03"] <- as.character("LS.I")
d$partition[d$Segment == "5A-04"] <- as.character("LS.I")
d$partition[d$Segment == "5B-06"] <- as.character("LS.I")
d$partition[d$Segment == "5B-07"] <- as.character("LS.I")
d$partition[d$Segment == "5B-09"] <- as.character("LS.I")

d$partition[d$Segment == "11A-01"] <- as.character("LS.I")
d$partition[d$Segment == "11A-02"] <- as.character("LS.I")
d$partition[d$Segment == "11A-03"] <- as.character("LS.I")
d$partition[d$Segment == "11A-05"] <- as.character("LS.I")
d$partition[d$Segment == "11A-06"] <- as.character("LS.I")
d$partition[d$Segment == "11B-03"] <- as.character("LS.I")
d$partition[d$Segment == "11B-04"] <- as.character("LS.I")
d$partition[d$Segment == "11B-06"] <- as.character("LS.I")
d$partition[d$Segment == "11B-07"] <- as.character("LS.I")
d$partition[d$Segment == "11B-08"] <- as.character("LS.I")

d$partition[d$Segment == "6A-01"] <- as.character("LS.II")
d$partition[d$Segment == "6A-09"] <- as.character("LS.II")
d$partition[d$Segment == "6B-01"] <- as.character("LS.II")
d$partition[d$Segment == "6B-09"] <- as.character("LS.II")

d$partition[d$Segment == "12A-04"] <- as.character("LS.II")
d$partition[d$Segment == "12A-05"] <- as.character("LS.II")
d$partition[d$Segment == "12A-06"] <- as.character("LS.II")
d$partition[d$Segment == "12B-03"] <- as.character("LS.II")
d$partition[d$Segment == "12B-04"] <- as.character("LS.II")
d$partition[d$Segment == "12B-05"] <- as.character("LS.II")

#LG
d$partition[d$Segment == "5A-02"] <- as.character("LG.I")
d$partition[d$Segment == "5A-05"] <- as.character("LG.I")
d$partition[d$Segment == "5A-06"] <- as.character("LG.I")
d$partition[d$Segment == "5A-07"] <- as.character("LG.I")
d$partition[d$Segment == "5A-08"] <- as.character("LG.I")
d$partition[d$Segment == "5A-09"] <- as.character("LG.I")
d$partition[d$Segment == "5B-01"] <- as.character("LG.I")
d$partition[d$Segment == "5B-02"] <- as.character("LG.I")
d$partition[d$Segment == "5B-03"] <- as.character("LG.I")
d$partition[d$Segment == "5B-04"] <- as.character("LG.I")
d$partition[d$Segment == "5B-05"] <- as.character("LG.I")
d$partition[d$Segment == "5B-08"] <- as.character("LG.I")

d$partition[d$Segment == "11A-04"] <- as.character("LG.I")
d$partition[d$Segment == "11B-05"] <- as.character("LG.I")

d$partition[d$Segment == "6A-02"] <- as.character("LG.II")
d$partition[d$Segment == "6A-03"] <- as.character("LG.II")
d$partition[d$Segment == "6A-04"] <- as.character("LG.II")
d$partition[d$Segment == "6A-05"] <- as.character("LG.II")
d$partition[d$Segment == "6A-06"] <- as.character("LG.II")
d$partition[d$Segment == "6A-07"] <- as.character("LG.II")
d$partition[d$Segment == "6A-08"] <- as.character("LG.II")

d$partition[d$Segment == "6B-02"] <- as.character("LG.II")
d$partition[d$Segment == "6B-03"] <- as.character("LG.II")
d$partition[d$Segment == "6B-04"] <- as.character("LG.II")
d$partition[d$Segment == "6B-05"] <- as.character("LG.II")
d$partition[d$Segment == "6B-06"] <- as.character("LG.II")
d$partition[d$Segment == "6B-07"] <- as.character("LG.II")
d$partition[d$Segment == "6B-08"] <- as.character("LG.II")

d$partition[d$Segment == "14A-01"] <- as.character("LG.II")
d$partition[d$Segment == "14B-10"] <- as.character("LG.II")# I added to take care of the 60 NA'S - Swapna

#UG
d$partition[d$Segment == "7A-01"] <- as.character("UG.I")
d$partition[d$Segment == "7A-02"] <- as.character("UG.I")
d$partition[d$Segment == "7A-03"] <- as.character("UG.I")
d$partition[d$Segment == "7B-01"] <- as.character("UG.I")
d$partition[d$Segment == "7B-02"] <- as.character("UG.I")
d$partition[d$Segment == "7B-03"] <- as.character("UG.I")

d$partition[d$Segment == "13A-01"] <- as.character("UG.I") #mistake
d$partition[d$Segment == "13A-02"] <- as.character("UG.I")
d$partition[d$Segment == "13A-03"] <- as.character("UG.I")
d$partition[d$Segment == "13A-05"] <- as.character("UG.I")
d$partition[d$Segment == "13A-06"] <- as.character("UG.I")
d$partition[d$Segment == "13A-07"] <- as.character("UG.I")
d$partition[d$Segment == "13A-08"] <- as.character("UG.I")
d$partition[d$Segment == "13B-01"] <- as.character("UG.I")
d$partition[d$Segment == "13B-02"] <- as.character("UG.I")
d$partition[d$Segment == "13B-03"] <- as.character("UG.I")
d$partition[d$Segment == "13B-04"] <- as.character("UG.I")
d$partition[d$Segment == "13B-06"] <- as.character("UG.I")
d$partition[d$Segment == "13B-07"] <- as.character("UG.I")
d$partition[d$Segment == "13B-08"] <- as.character("UG.I") #mistake

d$partition[d$Segment == "8A-01"] <- as.character("UG.II")
d$partition[d$Segment == "8A-02"] <- as.character("UG.II")
d$partition[d$Segment == "8A-03"] <- as.character("UG.II")
d$partition[d$Segment == "8A-04"] <- as.character("UG.II")
d$partition[d$Segment == "8B-01"] <- as.character("UG.II")
d$partition[d$Segment == "8B-02"] <- as.character("UG.II")
d$partition[d$Segment == "8B-03"] <- as.character("UG.II")
d$partition[d$Segment == "8B-04"] <- as.character("UG.II")

d$partition[d$Segment == "14A-02"] <- as.character("UG.II")
d$partition[d$Segment == "14A-03"] <- as.character("UG.II")
d$partition[d$Segment == "14A-04"] <- as.character("UG.II")
d$partition[d$Segment == "14A-05"] <- as.character("UG.II")
d$partition[d$Segment == "14A-06"] <- as.character("UG.II")
d$partition[d$Segment == "14A-07"] <- as.character("UG.II")
d$partition[d$Segment == "14A-08"] <- as.character("UG.II")
d$partition[d$Segment == "14B-01"] <- as.character("UG.II")
d$partition[d$Segment == "14B-02"] <- as.character("UG.II")
d$partition[d$Segment == "14B-03"] <- as.character("UG.II")
d$partition[d$Segment == "14B-04"] <- as.character("UG.II")
d$partition[d$Segment == "14B-05"] <- as.character("UG.II")
d$partition[d$Segment == "14B-06"] <- as.character("UG.II")
d$partition[d$Segment == "14B-07"] <- as.character("UG.II")
d$partition[d$Segment == "14B-08"] <- as.character("UG.II")
d$partition[d$Segment == "14B-09"] <- as.character("UG.II") #mistake
d$partition[d$Segment == "14A-10"] <- as.character("UG.II")# I added to take care of the 60 NA'S - Swapna
d$partition[d$Segment == "14A-09"] <- as.character("UG.II") # I added to take care of the 60 NA'S - Swapna

#MO
d$partition[d$Segment == "7A-04"] <- as.character("MO.I")
d$partition[d$Segment == "7A-05"] <- as.character("MO.I")
d$partition[d$Segment == "7B-04"] <- as.character("MO.I")
d$partition[d$Segment == "7B-05"] <- as.character("MO.I")

d$partition[d$Segment == "13A-04"] <- as.character("MO.I")
d$partition[d$Segment == "13B-05"] <- as.character("MO.I")

d$partition[d$Segment == "7A-06"] <- as.character("MO.II") #issue: split with MO.I
d$partition[d$Segment == "7A-07"] <- as.character("MO.II")
d$partition[d$Segment == "7B-06"] <- as.character("MO.II") #issue: split with MO.I
d$partition[d$Segment == "7B-07"] <- as.character("MO.II")  

d$partition[d$Segment == "8A-05"] <- as.character("MO.III")
d$partition[d$Segment == "8A-06"] <- as.character("MO.III")
d$partition[d$Segment == "8A-07"] <- as.character("MO.III") 
d$partition[d$Segment == "8B-05"] <- as.character("MO.III")
d$partition[d$Segment == "8B-06"] <- as.character("MO.III")
d$partition[d$Segment == "8B-07"] <- as.character("MO.III")

d$partition <- factor(d$partition)
levels(d$partition)
colnames(d)                        

```

#### save the tidied file
```{r}

write.csv(d, file = "leech_tidied_data.csv")
```

